CUNY SPS DATA-607 Week 1 Assignment

### 2018-08-28 DATA607/Week 1 Assignment/Week_1_Assignment.Rmd
More work to Week 1 Assignment

---
title: "CUNY SPS DATA-607 Week 1 Assignment"
subtitle: "Loading Data into a Data Frame"
author: "Mike Silva"
date: "August 27, 2018"
output: html_document
---

@@

```{r load-data}
df <- read.table('agaricus-lepiota.data', sep=',')
```

According to the documentation we should have 8,124 rows and 22 columns of data.

@@

df <- df %>%
  r
```

>>>>

---
title: "CUNY SPS DATA-607 Week 1 Assignment"
subtitle: "Loading Data into a Data Frame"
author: "Mike Silva"
date: "August 27, 2018"
output:
  html_document:
    keep_md: yes
---

@@

```{r load-data}
df <- read.table('agaricus-lepiota.data', sep=',', stringsAsFactors=FALSE)
head(df)
```

According to the documentation we should have 8,124 rows and 23 columns of data.

@@

df <- df %>%
  rename(type = V1,
         cap_shape = V2,
         cap_surface = V3,
         cap_color = V4,
         bruises = V5,
         odor = V6,
         gill_attachment = V7,
         gill_spacing = V8,
         gill_size = V9,
         gill_color = V10,
         stalk_shape = V11,
         stalk_root = V12,
         stalk_surface_above_ring = V13,
         stalk_surface_below_ring = V14,
         stalk_color_above_ring = V15,
         stalk_color_below_ring = V16,
         veil_type = V17,
         veil_color = V18,
         ring_number = V19,
         ring_type = V20,
         spore_print_color = V21,
         population = V22,
         habitat = V23)
```

## Step 4: Subsetting Data Frame

Before changing the data to be more meaningful I want to subset the data frame.  I would like to use this data to predict if a mushroom is poisonous or not.  According to the documentation the odor, spore print color, stalk surface below ring, and stalk color above ring are 4 attributes that can predict if a mushroom is poisonous with a 99.9% accuracy for these 23 species.

```{r subset-data}
df <- df %>%
  select(type, odor, spore_print_color, stalk_color_below_ring, stalk_color_above_ring)
```

Now that we have these subsets, let's examine the values:

```{r subset-tables}
lapply(df, function(x) table(x)) 
```

## Step 5: Recoding Variables

Now that we have the subset of columns we can recode the variables.  I will accomplish this with dplyr and recode statements.

```{r recode-variables}
df <- df %>%
  mutate(type = recode(type, p="Poisonous", e="Edible"),
         odor = recode(odor, a="Almond", c="Creosote", f="Foul", l="Anise", m="Musty", n="None", p="Pungent", s="Spicy", y="Fishy"),
         spore_print_color = recode(spore_print_color, b="Buff", h="Chocolate", k="Black", n="Brown", o="Orange", r="Green", u="Purple", w="White", y="Yellow"),
         stalk_color_below_ring = recode(stalk_color_below_ring, b="Buff", c="Cinnamon", e="Red", g="Gray", n="Brown", o="Orange", p="Pink", w="White", y="Yellow"),
         stalk_color_above_ring = recode(stalk_color_above_ring, b="Buff", c="Cinnamon", e="Red", g="Gray", n="Brown", o="Orange", p="Pink", w="White", y="Yellow"))
```

Now to examine the final product:

```{r final-product}
head(df)
tail(df)
```
